Operational Feature Selection in Gaussian Mixture Models

### 2016-09-26 paper.tex
update

        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
        \begin{prop}
        \label{eq:update-inv}
            (Forward update rule for inverse of covariance matrix)
            \begin{equation*}
                \fbox{$
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                A & v \\
                v^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
                $}
            \end{equation*}
        \end{prop}
        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} $.

@@

        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes
        \begin{prop}
            (Backward update rule for inverse of covariance matrix)
            \begin{equation*}
                \fbox{$(\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t$}
            \end{equation*}
        \end{prop}

        \begin{proof}
            \begin{align*}
                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
            \end{align*}
        \end{proof}

        A formula can also be deduced for the quadratical term of the decision function. Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained
        \begin{prop}
        \label{eq:update-quad}
            (Update rule for quadratical term)
            \begin{empheq}[box=\fbox]{align*}
                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
                &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{empheq}
        \end{prop}

        \begin{proof}
            \begin{align*}
                (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
                = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
                \left[\begin{array}{cc}
                \mathbf{A}   & \mathbf{v} \\
                \mathbf{v}^t & \frac{1}{\alpha}
                \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
                &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{align*}
        \end{proof}

        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
        \begin{prop}
        \label{eq:update-log}
            (Update rule for logdet)
            \begin{equation*}
                \fbox{$\log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha$}
            \end{equation*}
        \end{prop}

@@

% \appendices
% \section{Proof of the First Zonklar Equation}
% Appendix one text goes here.

>>>>

        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
        \begin{prop}
        \label{eq:update-inv}
            (Forward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                A & v \\
                v^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
            \end{equation*}
        \end{prop}
        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$.

@@

        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-inv-back}
            (Backward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t
            \end{equation*}
        \end{prop}

        A formula can also be deduced for the quadratical term of the decision function. Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-quad}
            (Update rule for quadratical term)
            \begin{align*}
                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
                &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{align*}
        \end{prop}

        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
        \begin{prop}
        \label{eq:update-log}
            (Update rule for logdet)
            \begin{equation*}
                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha
            \end{equation*}
        \end{prop}

@@

\appendices
\section{}
\label{app:proof-update}
\subsection{Proof of update rules}
    \subsubsection{Proposition~\ref{eq:update-inv-back}}
        \begin{proof}
            \begin{align*}
                \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
                &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
                &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
            \end{align*}
        \end{proof}

    \subsubsection{Proposition~\ref{eq:update-quad}}
        \begin{proof}
            \begin{align*}
                (&\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)}
                = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
                \left[\begin{array}{cc}
                \mathbf{A}   & \mathbf{v} \\
                \mathbf{v}^t & \frac{1}{\alpha}
                \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A} + x_k \mathbf{v}^t & (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} \end{array}\right]
                \left[\begin{array}{c} \mathbf{v} \\ x_k \end{array}\right] \\
                &= (\mathbf{x}^{(k-1)})^t \mathbf{A} \mathbf{x}^{(k-1)} + x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v} x_k + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}^{(k-1)})^{-1} + \alpha \mathbf{v} \mathbf{v}^t) \mathbf{x}^{(k-1)}\\
                &~~~+ 2 x_k \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha} \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} \mathbf{v}^t \mathbf{x}^{(k-1)} \\
                &~~~+ 2 \frac{x_k}{\alpha} \mathbf{v}^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha^2}) \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( (\mathbf{x}^{(k-1)})^t \mathbf{v} + \frac{x_k}{\alpha} )^2 \\
                &= (\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{align*}
        \end{proof}
